---
title: "README"
author: "Brian M. Schilder"
date: "Most recent update:<br> `r Sys.Date()`"
output: 
  md_document:
    variant: "markdown_github"
    toc: FALSE
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<!-- badges: start -->
<!-- badger::badge_codecov() --> 
<!-- badger::badge_last_commit()  -->
<!-- badger::badge_license() -->
[![](https://codecov.io/gh/RajLabMSSM/echotabix/branch/main/graph/badge.svg)](https://codecov.io/gh/RajLabMSSM/echotabix)
[![R-CMD-check](https://github.com/RajLabMSSM/echotabix/workflows/R-full/badge.svg)](https://github.com/RajLabMSSM/echotabix/actions)
[![](https://img.shields.io/github/last-commit/RajLabMSSM/echotabix.svg)](https://github.com/RajLabMSSM/echotabix/commits/main)
<!-- badges: end --> 


# echoverse: echotabix 

## Tabix indexing and querying.  

[`Tabix`](http://www.htslib.org/doc/tabix.html) is an extremely 
efficient CLI tool for querying large genomic data. 
However, using it reliably within R is still a challenge. 
`echotabix` efficiently creates and queries `Tabix` files across a variety of 
scenarios: 

- Checks whether files are already `Tabix`-ready (indexed), and if not, 
converts them.
- Queries files in tabular (e.g. *.csv*, *.tsv*, *.bed*) or VCF format.
- Queries files stored locally or on the remote server.
- Reports the size of the returned query.

 
```{r, eval=FALSE}
if(!"remotes" %in% rownames(installed.packages())){install.packages("remotes")}

remotes::install_github("RajLabMSSM/echotabix")
library(echotabix)
```


## [Documentation website](https://rajlabmssm.github.io/echotabix/)
## [Vignette](https://rajlabmssm.github.io/echotabix/articles/echotabix) 
<hr>

## Creator

<a href="https://bschilder.github.io/BMSchilder/" target="_blank">Brian
M. Schilder, Bioinformatician II</a>  
<a href="https://rajlab.org" target="_blank">Raj Lab</a>  
<a href="https://icahn.mssm.edu/about/departments/neuroscience" target="_blank">Department
of Neuroscience, Icahn School of Medicine at Mount Sinai</a>  
