---
title: "Getting Started"
author: "<h4>Authors: <i>`r auths <- eval(parse(text = gsub('person','c',read.dcf('../DESCRIPTION', fields = 'Authors@R'))));paste(auths[names(auths)=='given'],auths[names(auths)=='family'], collapse = ', ')`</i></h4>"
date: "<h4>Vignette updated: <i>`r format( Sys.Date(), '%b-%d-%Y')`</i></h4>"
output:
  BiocStyle::html_document
vignette: >
    %\VignetteIndexEntry{echotabix}
    %\usepackage[utf8]{inputenc}
    %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = TRUE)
has_internet <- curl::has_internet()
```

# Introduction

`echotabix` automates the creation and querying of
[tabix](http://www.htslib.org/doc/tabix.html)-indexed files.
It handles the full pipeline of liftover, sorting, compression, and indexing
with a single function call. Both local and remote tabix files
(VCF or tabular) can be queried, with automatic genome build detection
and liftover when needed.

# Installation

```{r, eval = FALSE}
if (!require("remotes")) install.packages("remotes")
remotes::install_github("RajLabMSSM/echotabix")
```

```{r}
library(echotabix)
```

# Convert

`convert()` takes a summary statistics file (or similar tabular genomic data)
and produces a sorted, bgzip-compressed, tabix-indexed file.

```{r convert_example, eval = has_internet}
## Download example full summary statistics (requires internet)
tmp <- echodata::example_fullSS()

tabix_files <- echotabix::convert(target_path = tmp,
                                  start_col = "BP")
```

# Query

`query()` automatically chooses the correct methods to perform a query
by inferring:

- Whether the file is local (e.g. on your computer)
or remote (e.g. on a server accessed with a URL).
- Whether the file is in VCF or table format.
- Whether your query's genome build (`query_genome`) matches
the full data's genome build (`target_genome`),
and if not performs liftover before querying.

`query_granges` is used to specify which coordinates you want to query from
the target file. This argument is flexible and
can take any of the following:

1. A `GRanges` object with one or more rows, across one or more chromosomes.
2. A `data.table` containing the columns "CHR", "POS", and "SNP",
which is automatically converted to a `GRanges` object.
3. The direct output of `echotabix::construct_query()`,
for additional flexibility and non-standard column names.

## Local file

```{r query_local, eval = has_internet}
query_dat <- echodata::BST1

query_res <- echotabix::query(target_path = tabix_files$path,
                              query_granges = query_dat)
knitr::kable(head(query_res))
```

## Customised queries

You can gain more customised control over the query
with the helper function `construct_query()` (which is called
internally by default when `query_granges` is not a `GRanges` object).

### Using min/max ranges

```{r query_ranges, eval = has_internet}
query_granges1 <- echotabix::construct_query(query_chrom = "chr4",
                                             query_start_pos = 5000,
                                             query_end_pos = 1e8L)
query_res2 <- echotabix::query(target_path = tabix_files$path,
                               query_granges = query_granges1)
```

### Using a query_dat object

```{r query_dat_custom, eval = has_internet}
query_dat2 <- echodata::LRRK2[1:100, ]
## Rename columns for demo purposes
data.table::setnames(query_dat2, c("CHR", "SNP"), c("seqnames", "rsID"))

query_granges2 <- echotabix::construct_query(query_dat = query_dat2,
                                             query_chrom_col = "seqnames",
                                             query_snp_col = "rsID")
query_res3 <- echotabix::query(target_path = tabix_files$path,
                               query_granges = query_granges2)
```

### Standardise column names

Instead of specifying each column name, you can also select
`standardise_colnames = TRUE` to automatically rename all columns to a
standard nomenclature using
[`MungeSumstats::format_header`
](https://neurogenomics.github.io/MungeSumstats/reference/standardise_header.html).

```{r query_standardise, eval = FALSE}
query_granges3 <- echotabix::construct_query(query_dat = query_dat2,
                                             standardise_colnames = TRUE)
query_res4 <- echotabix::query(target_path = tabix_files$path,
                               query_granges = query_granges3)
```

## Remote file

Next, we query whole-genome sequencing data from the
[1000 Genomes Project](https://www.internationalgenome.org/).
Specifically, we extract a given genomic range for a subset of samples
(individuals).

```{r query_remote, eval = has_internet}
target_path <- file.path(
    "ftp://ftp-trace.ncbi.nih.gov/1000genomes/ftp/release/20110521/",
    "ALL.chr4.phase1_release_v3.20101123.snps_indels_svs.genotypes.vcf.gz"
)
## Fewer samples will speed up queries substantially (15-30s vs. >1.2min).
samples <- c("HG00097", "HG00099", "HG00100", "HG00101", "HG00102")
dat2 <- echodata::BST1[1:50, ]

#### Query ####
vcf <- echotabix::query(target_path = target_path,
                        query_granges = dat2,
                        samples = samples,
                        query_save = TRUE,
                        as_datatable = FALSE,
                        query_force_new = TRUE)
print(vcf)
```

# Convert and query

`convert_and_query()` merges the `convert()` and `query()` functions
into a single pipeline.

- If the `target_path` file is not already a tabix file,
it will first be converted to a sorted, bgzip-compressed,
tabix-indexed file, and then proceed to the `query` step.
- If the `target_path` file is already a tabix file, the function will
detect this automatically and skip right to the `query` step.

```{r convert_and_query, eval = has_internet}
query_dat <- echodata::BST1
target_path2 <- echodata::example_fullSS()

query_res5 <- echotabix::convert_and_query(
    target_path = target_path2,
    target_start_col = "BP",
    query_granges = query_dat,
    query_force_new = TRUE)

knitr::kable(head(query_res5))
```

# Construct query (offline demo)

The `construct_query()` function works entirely offline using
bundled example datasets.

```{r construct_query_demo}
query_dat <- echodata::BST1
gr <- echotabix::construct_query(query_dat = query_dat)
print(gr)
```

```{r granges_to_string_demo}
echotabix::granges_to_string(gr = gr)
```

# Session Info

<details>

```{r Session Info}
utils::sessionInfo()
```

</details>
